### 2010-03-23 outline.tex

\documentclass[aps,prd,showpacs,groupedaddress,showkeys,preprintnumbers]{revtex4}
\input{packages}
\input{macros}

\begin{document}

\title{(OUTLINE DRAFT)\\ Low-latency search for gravitational-waves: prospects for near real time detection of compact binary coalescence}

\date{\today}

\author{Kipp Cannon}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Adrian Chapman}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Chad Hanna}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Drew Keppel}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Antony C. Searle}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\author{Alan J. Weinstein}
% \email{antony.searle@anu.edu.au}
\affiliation{The LIGO Laboratory - California Institute of Technology, Pasadena, CA 91125}

\begin{abstract}
The future of gravitational-wave astronomy permits the possibility that 
gravitational-wave detections from inspiralling compact objects could direct
electromagnetic followup searches. In order to be practical the latency of
gravitational-wave searches informing other astronomical instruments 
would have to be small to catch the peak electromagnetic radiation output. 
It is desirable to consider time domain filters for gravitational-wave data
to reduce the latency.   Typically the cost of time domain convolutions 
is prohibitive for any realizable analysis.  Here we present a method to
make the cost of time domain convolutions reasonable with the hope that it
could be employed for near real time detection of gravitational-wave signals
from compact binary mergers.  
\end{abstract}

\pacs{}

\keywords{Gravitational Waves, Laser Interferometry}

\preprint{}

\maketitle

\begin{enumerate}
\item{Introduction}
\begin{itemize}
\item{Introduce how GW searches could direct EM searches and do literature review}
\item{Introduce need for low latency searches and setup the discussion of our
method as well as a literature review for similar ideas}
\item{Outline the rest of the paper}
\end{itemize}
\item{Method}
\begin{itemize}
\item{Describe SVD decomposition of template banks}
\item{Describe selective downsampling of waveform portions}
\item{Describe computational cost / latency of analysis and compare to the 
standard method (theoretical)}
\end{itemize}
\item{Analysis}
\begin{itemize}
\item{Describe the input to the simulations we will do, i.e. the template bank,
the match conditions, the whitening, all the gory details...}
\end{itemize}
\item{Results}
\begin{itemize}
\item{Show results of our ability to reconstruct waveforms and compare to 
``theory"}
\item{Show results of the latency and computational costs and compare to ``theory"}
\end{itemize}
\item{Conclusions}
\begin{itemize}
\item{Sum up the status and describe the future work}
\end{itemize}
\end{enumerate}
\begin{acknowledgments}
LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the National Science Foundation and operates under cooperative agreement PHY-0107417. This paper has LIGO Document Number LIGO-???
\end{acknowledgments}

\end{document}