### 2011-05-27 figures/diagram.tex
Added diagram
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@763 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass{amsart}
\usepackage[usenames]{color}
\usepackage{tikz}
\usepackage{signalflowdiagram}
\usetikzlibrary{fit,calc,chains,scopes}
\begin{document}

\tikzstyle{junction}=[draw, fill, circle, line width=0pt, inner sep=0pt, minimum size=3pt]
\tikzstyle{signal}=[draw, fill=white, circle, inner sep=0pt, minimum size=4pt]
\tikzstyle{resample}=[draw, circle, line width=1pt, inner sep=0pt, minimum size=0.3cm]

\begin{figure}
	\begin{center}
		\begin{tikzpicture}[
			node distance=2mm,
			every path/.style={line width=1pt},
			every node/.style={line width=1pt}
		]
			\foreach \x in {0,1,2} {
				\node[draw] at (0, {\x == 0 ? 0cm : (\x == 1 ? -3cm : -7.5cm)}) (u0) {};
				\node[draw, below=of u0] (u1) {};
				\node[below=of u1] (uellip\x) {$\vdots$};
				\node[draw, below=of uellip\x] (u2) {};

				\node[junction] at ($0.5*(u0) + 0.5*(u2) - (0.5, 0)$) (in\x tee) {};
				\node[resample, left=1.5cm of in\x tee] (in\x) {\tiny$\downarrow$};

				\foreach \i in {0,1,2} {
					\draw (in\x tee) |- (u\i);
					\node[junction, right=1cm of u\i] (tee u\i) {};
					\draw (u\i) -- (tee u\i);
				}

				\node[adder, right=2cm of tee u0] (v0\x) {};
				\node[adder, below=of v0\x] (v1\x) {};
				\node[adder, below=of v1\x] (v2\x) {};
				\node[right=0.85cm of uellip\x] {$\vdots$};
				\node[adder, right=2cm of tee u2] (v3\x) {};
				\node[above=0cm of v3\x] {$\vdots$};

				\foreach \i in {0,1,2}
					\foreach \j in {0,1,2,3}
						\draw (tee u\i) -- (v\j\x);

				\node[draw,dashed,line width=0.5pt,fit=(in\x tee) (u0) (u2)] (ortho\x) {};
				\node[draw,dashed,line width=0.5pt,fit=(tee u0) (tee u2) (v0\x) (v3\x)] (reconstruct\x) {};
			}

			\foreach \x in {1,2} {
				\node[resample, right=1cm of v0\x, yshift=0.5cm] (resample0\x) {\tiny$\uparrow$};
				\node[resample, right=1.5cm of v0\x, yshift=0.5cm] (resample1\x) {\tiny$\uparrow$};
				\node[resample, right=2cm of v0\x, yshift=0.5cm] (resample2\x) {\tiny$\uparrow$};
				\node[right=2.5cm of v0\x, yshift=0.5cm] (dots) {$\cdots$};
				\node[below=of dots] (dots) {$\cdots$};
				\node[below=of dots] (dots) {$\cdots$};
				\node[resample, right=3.5cm of v0\x, yshift=0.5cm] (resample3\x) {\tiny$\uparrow$};
			}

			\foreach \x in {0,1} {
				\node[adder, right=1cm of v0\x] (adder0\x) {};
				\node[adder, right=1.5cm of v1\x] (adder1\x) {};
				\node[adder, right=2cm of v2\x] (adder2\x) {};
				\node[adder, right=3.5cm of v3\x] (adder3\x) {};
				\foreach \i in {0,1,2,3} {
					\draw[->] (v\i \x) -- (adder\i \x);
				}
			}

			\foreach \i in {0,1,2,3} {
				\node[signal, right=4.5cm of v\i 0, label=right:$\rho_\i$] (out\i) {};
				\draw[->] (v\i 2) -| (resample\i 2);
				\draw[->] (resample\i 2) -- (adder\i 1) node[above=0.5cm of resample\i 2,fill=white] {$\vdots$};
				\draw[->] (adder\i 1) -- (resample\i 1);
				\draw[->] (resample\i 1) -- (adder\i 0);
				\draw[->] (adder\i 0) -- (out\i);
			}

			\node[draw, left=0.5cm of in0] (whiten) {\emph{whiten}};
			\node[signal, label=left:$x$, label=below:{\small 16384 Hz}, left=0.75cm of whiten] (x) {};
			\node[below=1.5cm of in1] (ellip) {$\vdots$};
			\node[right=1.6cm of ellip] {$\vdots$};
			\node[right=4.1cm of ellip] {$\vdots$};
			\node[above=1mm of out3] {\vdots};
			\draw[->] (x) -- (whiten);
			\draw[->] (whiten) -- (in0);
			\draw[->] (in0) -- (in1);
			\draw (in1) -- (ellip);
			\draw[->] (ellip) -- (in2);

			\draw (in0) -- (in0tee) node[midway,above] {\small 4096 Hz};
			\draw (in1) -- (in1tee) node[midway,above] {\small 512 Hz};
			\draw (in2) -- (in2tee) node[midway,above] {\small 32 Hz};

			\node[above=0.5cm of ortho0,align=center] (fir label) {Orthogonal \\ FIR filters \\ $\overbrace{\hspace{1cm}}^{}$};
			\node[right=0.1cm of fir label,align=center] {Reconstruction \\ matrices \\ $\overbrace{\hspace{2.75cm}}^{}$};
			\node[right=3.5cm of fir label,align=center] {Upsampling and \\ SNR accumulation \\ $\overbrace{\hspace{2.75cm}}^{}$};
			\node[left=0.1cm of fir label,align=center] {Down-\\ sampling \\ $\overbrace{\hspace{1cm}}^{}$};
			\node[left=1.75cm of fir label,align=center] {Data acquisition \\ and whitening \\ $\overbrace{\hspace{2cm}}^{}$};

		\end{tikzpicture}
		\caption{Schematic of LLOID pipeline illustrating signal flow.  Circles with arrows represent upsampling \protect\tikz{\protect \node[resample] {\tiny$\uparrow$};} or downsampling \protect\tikz{\protect \node[resample] {\tiny$\downarrow$};}.  Circles with plus signs represent summing junctions \protect\tikz{\protect \node[line width=1pt,adder] {\tiny$\uparrow$};}.  Squares \protect\tikz{\protect \node[line width=1pt,draw] {};} stand for FIR filters.  Sample rate decreases from the top of the diagram to the bottom.}
	\end{center}
\end{figure}

This diagram is a work in progress.  Some corrections to make:

{\color{red}
\begin{itemize}
\item Add delay lines.
\item Switch first downsampler and whitener.
\item Resamplers are connected in series, not parallel.
\item Translate figure to the left to align it visually with the caption.
\end{itemize}
}

\end{document}