### 2011-05-30 figures/diagram.tex
Removed whitener
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@776 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass{amsart}
\usepackage{tikz}
\usepackage{signalflowdiagram}
\usetikzlibrary{external,fit,calc,chains,scopes}
\tikzexternalize[force remake=true]
\begin{document}

\tikzstyle{junction}=[draw, fill, circle, line width=0pt, inner sep=0pt, minimum size=3pt]
\tikzstyle{signal}=[draw, fill=white, circle, inner sep=0pt, minimum size=4pt]
\tikzstyle{resample}=[draw, circle, line width=1pt, inner sep=0pt, minimum size=0.3cm]

\tikzsetnextfilename{lloid-diagram}
\begin{tikzpicture}[
	node distance=2mm,
	every path/.style={line width=1pt},
	every node/.style={line width=1pt}
]
	\foreach \x in {0,1,2} {
		\node[draw] at (0, {\x == 0 ? 0cm : (\x == 1 ? -3cm : -7.5cm)}) (u0) {};
		\node[draw, below=of u0] (u1) {};
		\node[below=of u1] (uellip\x) {$\vdots$};
		\node[draw, below=of uellip\x] (u2) {};

		\node[junction] at ($0.5*(u0) + 0.5*(u2) - (0.5, 0)$) (in\x tee) {};

		\foreach \i in {0,1,2} {
			\draw (in\x tee) |- (u\i);
			\node[junction, right=1cm of u\i] (tee u\i) {};
			\draw (u\i) -- (tee u\i);
		}

		\node[adder, right=2cm of tee u0] (v0\x) {};
		\node[adder, below=of v0\x] (v1\x) {};
		\node[adder, below=of v1\x] (v2\x) {};
		\node[right=0.85cm of uellip\x] {$\vdots$};
		\node[adder, right=2cm of tee u2] (v3\x) {};
		\node[above=0cm of v3\x] {$\vdots$};

		\foreach \i in {0,1,2}
			\foreach \j in {0,1,2,3}
				\draw (tee u\i) -- (v\j\x);

		\node[draw,dashed,line width=0.5pt,fit=(in\x tee) (u0) (u2)] (ortho\x) {};
		\node[draw,dashed,line width=0.5pt,fit=(tee u0) (tee u2) (v0\x) (v3\x)] (reconstruct\x) {};
	}

	\foreach \x in {1,2} {
		\node[resample, right=1cm of v0\x, yshift=0.5cm] (resample0\x) {\tiny$\uparrow$};
		\node[resample, right=1.5cm of v0\x, yshift=0.5cm] (resample1\x) {\tiny$\uparrow$};
		\node[resample, right=2cm of v0\x, yshift=0.5cm] (resample2\x) {\tiny$\uparrow$};
		\node[right=2.5cm of v0\x, yshift=0.5cm] (dots) {$\cdots$};
		\node[below=of dots] (dots) {$\cdots$};
		\node[below=of dots] (dots) {$\cdots$};
		\node[resample, right=3.5cm of v0\x, yshift=0.5cm] (resample3\x) {\tiny$\uparrow$};
	}

	\node[junction, left=1.9cm of in0tee] (in0) {};
	\node[resample, left=1.75cm of in1tee] (in1) {\tiny$\downarrow$};
	\node[resample, left=1.75cm of in2tee] (in2) {\tiny$\downarrow$};

	\foreach \x in {0,1} {
		\node[adder, right=1cm of v0\x] (adder0\x) {};
		\node[adder, right=1.5cm of v1\x] (adder1\x) {};
		\node[adder, right=2cm of v2\x] (adder2\x) {};
		\node[adder, right=3.5cm of v3\x] (adder3\x) {};
		\foreach \i in {0,1,2,3} {
			\draw[->] (v\i \x) -- (adder\i \x);
		}
	}

	\foreach \i in {0,1,2,3} {
		\node[signal, right=4.5cm of v\i 0, label=right:$\rho_\i$] (out\i) {};
		\draw[->] (v\i 2) -| (resample\i 2);
		\draw[->] (resample\i 2) -- (adder\i 1) node[above=0.5cm of resample\i 2,fill=white] {$\vdots$};
		\draw[->] (adder\i 1) -- (resample\i 1);
		\draw[->] (resample\i 1) -- (adder\i 0);
		\draw[->] (adder\i 0) -- (out\i);
	}

	\node[signal, label=left:$x$, left=0.75cm of in0] (x) {};
	%\node[draw, left=0.5cm of in0] (whiten) {\emph{whiten}};
	%\node[signal, label=left:$x$, label=below:{\small 16384 Hz}, left=0.75cm of whiten] (x) {};
	\node[below=1.5cm of in1] (ellip) {$\vdots$};
	\node[right=1.6cm of ellip] {$\vdots$};
	\node[right=4.1cm of ellip] {$\vdots$};
	\node[above=1mm of out3] {\vdots};
	\draw[->] (x) -- (in0);
	%\draw[->] (x) -- (whiten);
	%\draw[->] (whiten) -- (in0);
	\draw[->] (in0) -- (in1);
	\draw (in1) -- (ellip);
	\draw[->] (ellip) -- (in2);

	\draw (in0) -- (in0tee) node[midway,above] {\small 4096 Hz};
	\draw (in1) -- (in1tee) node[midway,draw,fill=white,label=above:{\small delay}, label=below:{\small 512 Hz}] {$z^{-t_1 f_1}$};
	\draw (in2) -- (in2tee) node[midway,draw,fill=white,label=above:{\small delay},label=below:{\small 32 Hz}] {$z^{-\cdots}$};

	\node[above=0.5cm of ortho0,align=center] (fir label) {Orthogonal \\ FIR filters \\ $\overbrace{\hspace{1cm}}^{}$};
	\node[right=0.1cm of fir label,align=center] {Reconstruction \\ matrices \\ $\overbrace{\hspace{2.75cm}}^{}$};
	\node[right=3.5cm of fir label,align=center] {Interpolation and \\ SNR accumulation \\ $\overbrace{\hspace{2.75cm}}^{}$};
	\node[left=0.3cm of fir label,align=center] {Decimation \\ of input \\ $\overbrace{\hspace{1cm}}^{}$};
	%\node[left=1.75cm of fir label,align=center] {Data acquisition \\ and whitening \\ $\overbrace{\hspace{2cm}}^{}$};

\end{tikzpicture}

\tikzsetnextfilename{upsample-symbol}
\begin{tikzpicture}
	\node[resample] {\tiny$\uparrow$};
\end{tikzpicture}

\tikzsetnextfilename{downsample-symbol}
\begin{tikzpicture}
	\node[resample] {\tiny$\downarrow$};
\end{tikzpicture}

\tikzsetnextfilename{adder-symbol}
\begin{tikzpicture}
	\node[line width=1pt,adder] {};
\end{tikzpicture}

\tikzsetnextfilename{fir-symbol}
\begin{tikzpicture}
	\node[line width=1pt,draw] {};
\end{tikzpicture}

\end{document}