### 2011-06-01 analysis.tex
more stuff
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@801 ccdd05d5-56c1-461a-bf98-85f6876523b4

\section{Implementation}

In this section we describe an implementation of the \lloid\ method described
in section \ref{label{sec:method}} suitable for rapid gravitational-wave
searches for compact binary coalescence.  The \lloid\ method requires several
computations that can be done before the analysis is actually underway.  Thus
we divide the procedure into two stages 1) an offline planning stage and 2) an
online, low-latency filtering stage.  The offline stage can be done before the
analysis is started and updated asychronously, whereas the online stage must
keep up with the detector output and produce search results as rapidly as
possible.  In the next two subsections we describe what these stages entail.

\subsection{Planning stage}

The choice of filter waveforms and singular value decomposition can be done in
advance and will be valid as long as the detector noise spectrum remains
roughly constant.  New filter waveforms can be computed asynchronously using
updated spectrum estimates as they are available.

The planning stage begins with choosing templates that cover the space of mass
parameters with a hexagonal grid \cite{PhysRevD.76.102004} in order to satisfy
the minimum match criterion, which assures a user specified maximum loss in SNR
for signals that fall in-between the chosen templates.  Next, the templates are
subdivided into groups of neighbors called ``sub-banks'' that are appropriately
sized so that each bank can be efficiently handled by a single computer.
Dividing the mass space into smaller sub-banks also aids in the computational
cost of the singular value decomposition and is the approach considered
in~\cite{Cannon:2010p10398}.  Using our understanding of the time-frequency
evolution of the templates, we choose time slice boundaries as in \eqref{FIXME}
such that all of the templates within a sub-bank are sub-critically sampled at
progressively lower sample rates.  Next, the templates within the sub-bank are
realized as \textsc{fir} filter coefficients.  For each time slice, the
templates are downsampled to the appropriate sample rate.  Finally, the
\textsc{svd} is applied to each time slice in the sub-bank in order to produce
a set of orthogonal \textsc{fir} filters and a reconstruction matrix that maps
them back to the original templates as described in \eqref{FIXME}.  The
downsampled orthogonal \textsc{fir} filter coefficients, the reconstruction
matrix, and the time slice boundaries are all saved to disk.

\subsection{Filtering stage}

The \lloid\ algorithm could be used in a true sample-in-sample-out realtime
system.  However, such a system would likely require integration directly into
the data acquisition and storage system of the gravitational wave
observatories.  A slightly more modest goal is to leverage existing low
latency, but not realtime, signal processing infrastructure in order to
implement the \lloid\ algorithm.  For the near-term this should be a viable 
solution for ultra-low latency searches with $\sim 1s$ intrinsic latency.

We have implemented a prototype of the low latency filtering stage using an
open source signal processing environment called \gstreamer\ \cite{gstreamer}.
\gstreamer\ is a vital component of many Linux systems, providing media
playback, authoring, and streaming on devices from cell phones to desktop
computers to streaming media servers.  Given the similarities of gravitational
wave detector data to audio data it is not surprising that \gstreamer\ is
useful for our purpose. \gstreamer\ also provides some useful stock signal
processing elements such as decimators, filters, and interpolators.  
We have extended the \gstreamer\ framework by developing custom elements
that are available in the \gstlal\ library~\cite{gstlal}.
Figure \ref{fig:pipeline} describes schematically how we implement the \lloid\
algorithm using \gstlal\ and \gstreamer\ components.
%
%
\begin{figure}[htbp]
	\includegraphics{figures/lloid-diagram.pdf}
	\caption{\label{fig:pipeline} Schematic of LLOID pipeline illustrating
signal flow.  Circles with arrows represent upsampling
\protect\includegraphics{figures/upsample-symbol.pdf} or downsampling
\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
signs represent summing junctions
\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
\protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.  Sample
rate decreases from the top of the diagram to the bottom.}
\end{figure}

The filter pipeline drawn in figure \ref{fig:pipeline} consists of several
distinct stages and parallel branches.  We outline the stages below.

\subsubsection{Decimation}

First, the sample rate of the whitened detector data is reduced to successively
lower sample rates by decimation.  Decimation involves applying an antialiasing
filter to the data, and then downsampling by deleting samples.  We use a
192-tap \textsc{fir} decimator provided by {\tt Gstreamer's audioresample}
element.  The detector data is provided at every power-of-two sample rate
required by the template time sliced described in \eqref{FIXME}.  These
parallel decimated data streams are fed into parallel fir filter banks in the
next stage.

\subsubsection{Orthogonal \textsc{fir} filters}

The \fir\ filter banks are implemented using a \gstlal\ element called {\tt
lal\_firbank}, which produces N channels of filter output from an NxM matrix of
\fir\ filter coefficients.  This element is used in parallel branches in the
pipeline to implement the filtering of the SVD basis filters in each time
slice.  Rather than implement the time sliced templates as zero padded \fir\
filters as described in \eqref{FIXME} we instead implement them as shorter, but
appropriately time advanced \fir\ filters that contain only the nonzero
samples.  Adding the appropriate time offset to the filter output makes up for
the lack of explicit zero padding.  The orthogonal filter outputs must be
reconstructed into the filter output for the underlying physical filter ouputs.

\subsubsection{Reconstruction}

The orthogonal filter output provided by the SVD must be used in linear
combinations in order to obtain the physical filter output for the template
waveforms.  This involves a matrix multiplication.  We use a \gstlal\ element
called {\tt lal\_matrix\_mixer} on the output of each orthogonal filter bank in
order to produce the filter output for the original waveforms in each time
slice.  At this point the time slice filter outputs are not yet at the same
sample rate.

\subsubsection{Interpolation}

Before we can construct the final \SNR\, the individual time slice filter
outputs must be brought to a common sample rate, or upsampled.  This is done
using the same \gstreamer\ element as was used for decimation, {\tt
audioresample}.

\subsubsection{\SNR accumulation}

In order to accumulate \SNR\ for an early warning, the time slice filter
output is upsampled to the next highest frequency time slice and summed. This
process is iterated until the final \SNR is computed.  However at each stage
of upsampling the summed filter outputs are related to the early warning \SNR
by a simple normalization factor that can be computed in advance. Thus the
\lloid\ algorithm and this implementation leads to a simple early warning
pipeline with little additional work.