### 2011-06-09 inspiral_svd.tex
Converted to AASTeX
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@997 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass[preprint2]{aastex}
\input{packages}
\input{macros}

\begin{document}

\title{Technique for \earlywarning\ detection of compact binary coalescence
and its implications for multi-messenger astronomy}

\shorttitle{Early-warning of compact binary coalescence}

\author{
	Kipp Cannon\altaffilmark{1},
	Romain Cariou\altaffilmark{2},
	Adrian Chapman\altaffilmark{3},
	Mireia Crisp\'{i}n-Ortuzar\altaffilmark{4},
	Nickolas Fotopoulos\altaffilmark{3},
	Melissa Frei\altaffilmark{5},
	Chad Hanna\altaffilmark{6},
	Erin Kara\altaffilmark{7},
	Drew Keppel\altaffilmark{8,9},
	Laura Liao\altaffilmark{10},
	Stephen Privitera\altaffilmark{3},
	Antony Searle\altaffilmark{3},
	Leo Singer\altaffilmark{3}, and
	Alan Weinstein\altaffilmark{3}
}

\altaffiltext{1}{Canadian Institute for Theoretical Astrophysics, Toronto, ON, Canada}
\altaffiltext{2}{D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
\altaffiltext{3}{LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA} 
\altaffiltext{4}{Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot, Spain} 
\altaffiltext{5}{University of Texas at Austin, Austin, TX, USA}
\altaffiltext{6}{Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
\altaffiltext{7}{Department of Physics and Astronomy, Barnard College, Columbia University, New York, NY, USA}
\altaffiltext{8}{Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
\altaffiltext{9}{Leibniz Universit\"{a}t Hannover, Hannover, Germany}
\altaffiltext{10}{Ryerson University, Toronto, ON, Canada}

\keywords{gamma ray burst: general --- gravitational waves --- methods: data analysis --- methods: numerical}

\begin{abstract}
The rapid detection of compact binary coalescence with a network of advanced
gravitational-wave detectors will offer a unique opportunity for
multi-messenger astronomy.  Prompt detection alerts to the astronomical
community may make it possible to observe the onset of electromagnetic emission
from compact binary coalescence.  We demonstrate a computationally practical
analysis strategy that produces early-warning triggers even before
gravitational radiation from the final merger has arrived at the detectors.
With current rate estimates for the Advanced \LIGO{} design configuration, it
should be possible to detect $\sim$10 such sources earlier than 10 seconds
before merger in 1 year of live time using the method we describe.
\end{abstract}

\input{introduction}
\input{method}
\input{implementation}
\input{results}
\input{conclusions}

\acknowledgements

\LIGO\ was constructed by the California Institute of Technology and
Massachusetts Institute of Technology with funding from the National Science
Foundation and operates under cooperative agreement
\textsc{phy}-\oldstylenums{0107417}.  CH would like to thank Ilya Mandel for
many discussions about rate estimates and the prospects of early detection.  CH
would also like to thank Patrick Brady for countless fruitful conversations
about low latency analysis methods.  NF would like to thank Alessandra Corsi
for illuminating discussions on astronomical motivations.  DK was supported
from the Max Planck Gesellschaft. This research is supported by the National
Science Foundation through a Graduate Research Fellowship to LS and by a
fellowship from The Perimeter Institute for Theoretical Physics to CH.

This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}.

\bibliographystyle{apj}
\bibliography{references}

\end{document}