### 2011-06-29 inspiral_svd.tex
mireia asked to remove accnt
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@1197 ccdd05d5-56c1-461a-bf98-85f6876523b4

\documentclass[preprint2]{aastex}
\input{packages}
\input{macros}

\begin{document}

\title{Technique for \earlywarning\ filtering for compact binary coalescence
and its implications for multi-messenger astronomy}

\shorttitle{Early-warning detection of \GW\ inspirals}

\author{
	Kipp Cannon\altaffilmark{1},
	Romain Cariou\altaffilmark{2},
	Adrian Chapman\altaffilmark{3},
	Mireia Crispin-Ortuzar\altaffilmark{4},
	Nickolas Fotopoulos\altaffilmark{3},
	Melissa Frei\altaffilmark{5},
	Chad Hanna\altaffilmark{6},
	Erin Kara\altaffilmark{7},
	Drew Keppel\altaffilmark{8,9},
	Laura Liao\altaffilmark{10},
	Stephen Privitera\altaffilmark{3},
	Antony Searle\altaffilmark{3},
	Leo Singer\altaffilmark{3}, and
	Alan Weinstein\altaffilmark{3}
}
% we either need to include all of our emails, or none.
%\email{leo.singer@ligo.org}

\altaffiltext{1}{Canadian Institute for Theoretical Astrophysics, Toronto, ON, Canada}
\altaffiltext{2}{D\'{e}partement de physique, \'{E}cole Normale Sup\'{e}rieure de Cachan, Cachan, France}
\altaffiltext{3}{LIGO Laboratory - California Institute of Technology, Pasadena, CA, USA} 
\altaffiltext{4}{Facultat de F\'{i}sica, Universitat de Val\`{e}ncia, Burjassot, Spain} 
\altaffiltext{5}{University of Texas at Austin, Austin, TX, USA}
\altaffiltext{6}{Perimeter Institute for Theoretical Physics, Waterloo, ON, Canada}
\altaffiltext{7}{Department of Physics and Astronomy, Barnard College, Columbia University, New York, NY, USA}
\altaffiltext{8}{Albert-Einstein-Institut, Max-Planck-Institut f\"{u}r Gravitationphysik, Hannover, Germany}
\altaffiltext{9}{Leibniz Universit\"{a}t Hannover, Hannover, Germany}
\altaffiltext{10}{Ryerson University, Toronto, ON, Canada}

\keywords{gamma ray burst: general --- gravitational waves --- methods: data analysis --- methods: numerical}

\begin{abstract}
Rapid detection of compact binary coalescence with a network of advanced
gravitational-wave detectors will offer a unique opportunity for
multi-messenger astronomy.  Prompt detection alerts to the astronomical
community may make it possible to observe the onset of electromagnetic emission
from compact binary coalescence.  We demonstrate a computationally practical
filtering strategy that could produce early-warning triggers before
gravitational radiation from the final merger has arrived at the detectors.
\end{abstract}

\input{introduction}
\input{method}
\input{implementation}
\input{results}
\input{conclusions}

\acknowledgements

\LIGO\ was constructed by the California Institute of Technology and
Massachusetts Institute of Technology with funding from the National Science
Foundation and operates under cooperative agreement
\textsc{phy}-\oldstylenums{0107417}.  CH would like to thank Ilya Mandel for
many discussions about rate estimates and the prospects of early detection.  CH
would like to thank Patrick Brady for countless fruitful conversations about
low latency analysis methods.  CH would like to thank John Zweizig for
disscussions about \LIGO\ data acquisition.  NF would like to thank Alessandra
Corsi for illuminating discussions on astronomical motivations.  This research
is supported by the National Science Foundation through a Graduate Research
Fellowship to LS and by the Perimeter Institute for Theoretical Physics through
a fellowship to CH. DK is supported from the Max Planck Gesellschaft.

This paper has \LIGO\ Document Number \LIGO-\textsc{p}\oldstylenums{0900004}-v\oldstylenums{21}.

\bibliographystyle{apj}
\bibliography{references}

\end{document}